
---
output: 
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 9: Visualization - Revealing a graph"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment9.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```


-------

**Notice:** *In this assignment, all the plot should have title, caption, and axis labels. *

1. Use the dataset by of covid 19 by WHO at https://covid19.who.int/WHO-COVID-19-global-data.csv. Find the three countries with the most numbers of deaths by Covid-19. 

Hint: 

-  Use `filter` to select the most current date 

-  Pipe with `arrange(-Cumulative_deaths)` to rank the data by the number of cumulative deaths
```{r}
library(gganimate)
library(ggplot2)
library(tidyverse)
library(lubridate)
setwd('C:/Users/student/Desktop/New folder/math_421')
df <- read_csv('WHO-COVID-19-global-data.csv')
```
```{r}
df <- df %>%
  mutate(Date_reported = as.Date(Date_reported, format = "%Y-%m-%d"))

latest_date <- max(df$Date_reported, na.rm = TRUE)
```
```{r}
top_countries_deaths <- df %>%
  filter(Date_reported == latest_date) %>%
  arrange(desc(Cumulative_deaths)) %>%
  slice(1:3) %>%
  select(Country, Cumulative_deaths)

top_countries_deaths
```

2. Use `transition_reveal` to make a plot revealing the number of deaths in the three countries with the most numbers of deaths
```{r}
top_countries_data <- df %>%
  filter(Country %in% top_countries_deaths$Country)
```
```{r}
death_plot <- ggplot(top_countries_data, aes(x = Date_reported, y = Cumulative_deaths, color = Country)) +
  geom_line(size = 1) +
  labs(
    title = "Cumulative Deaths Over Time",
    x = "Date",
    y = "Cumulative Deaths",
    caption = "Data Source: WHO"
  ) +
  theme_minimal() +
  transition_reveal(Date_reported)

animate(death_plot, nframes = 100, fps = 10)
```

3. Create the new variable (column) `death_per_cases` recording the number of deaths per cases (Hint: divide cumulative deaths by cumulative cases).  What are the three countries with the highest deaths per cases? 
```{r}
df <- df %>%
  mutate(death_per_cases = Cumulative_deaths / Cumulative_cases)
```
```{r}
top_countries_dpc <- df %>%
  filter(Date_reported == latest_date) %>%
  arrange(desc(death_per_cases)) %>%
  slice(1:3) %>%
  select(Country, death_per_cases)

top_countries_dpc
```


4. Use `transition_reveal` to make a plot revealing the number of deaths per cases of the US, Italy and Mexico. 
```{r}
selected_countries <- df %>%
  filter(Country %in% c("United States of America", "Italy", "Mexico"))
```
```{r}
dpc_plot <- ggplot(selected_countries, aes(x = Date_reported, y = death_per_cases, color = Country)) +
  geom_line(size = 1) +
  labs(
    title = "Deaths Per Cases Over Time",
    x = "Date",
    y = "Deaths Per Cases",
    caption = "Data Source: WHO"
  ) +
  theme_minimal() +
  transition_reveal(Date_reported)

animate(dpc_plot, nframes = 100, fps = 10)
```


5. Import the following data https://raw.githubusercontent.com/owid/covid-19-data/master/public/data/vaccinations/vaccinations.csv

Use `transition_reveal` to plot the total vaccinations per hundred by level of income. Hint: the levels of income are identified in the `location` variable. Put a short comment on the caption of the plot. 
```{r}
setwd('C:/Users/student/Desktop/New folder/math_421')
df <- read_csv("vaccinations.csv")
```
```{r}
vaccination_summary <- df %>%
  group_by(location) %>%
  summarize(
    max_vaccinations_per_hundred = max(people_vaccinated_per_hundred, na.rm = TRUE)
  )
```
```{r}
vaccination_plot <- ggplot(vaccination_summary, aes(x = reorder(location, -max_vaccinations_per_hundred), y = max_vaccinations_per_hundred, fill = location)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Vaccination Rates by Location",
    x = "Location",
    y = "Max Vaccinations Per Hundred",
    caption = "Data Source: Our World in Data"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
```{r}
print(vaccination_plot)
```

                      

