---
title: "Comprehensive Flex Dashboard"
author: "Liz Waddington"
date: "2024-12-13"
output: html_document
runtime: shiny
---

```{r}
library(shiny)
library(dplyr)
library(ggplot2)
library(DT)
library(plotly)
library(flexdashboard)
```
Introduction {data-height=200}
-------------------------------------
{.sidebar}
=======================================================================

```{r}
# Introduction text
includeText <- "
This dashboard analyzes the Gapminder dataset to explore global trends across countries 
in life expectancy, GDP, and population. Users can filter data using various inputs 
and interact with visualizations. 
"

div(style = "padding: 10px; font-size: 16px;", 
    includeText)
```
```{r}
library("gapminder")
data("gapminder") 
head(gapminder) 

img_src <- "Gapminder_Foundationr_Logo.png"
tags$img(src = img_src, height = "100px")
```

{data-width=300}
-------------------------------------
```{r}
selectInput("continent", 
            "Select Continent:", 
            choices = unique(gapminder$continent), 
            selected = "Asia")

sliderInput("year_range", 
            "Year Range:", 
            min = min(gapminder$year), 
            max = max(gapminder$year), 
            value = c(1960, 2005))

selectInput("metric", 
            "Select Metric:", 
            choices = c("Life Expectancy" = "lifeExp", 
                        "GDP per Capita" = "gdpPercap", 
                        "Population" = "pop"), 
            selected = "lifeExp")

checkboxGroupInput("country", 
                   "Choose Countries:", 
                   choices = unique(gapminder$country),
                   selected = c("India", "United States"))

radioButtons("highlight", 
             "Highlight Region:",
             choices = unique(gapminder$continent))
```

Main Tab 1
=======================================================================

Column {data-width=700}
-------------------------------------
```{r}
renderDataTable({
  gapminder %>%
    filter(
      continent %in% input$continent,
      year >= input$year_range[1] & year <= input$year_range[2],
      country %in% input$country
    ) %>%
    datatable()
})
```

Main Tab 2
=======================================================================

Tab 2: Trends Over Time {data-height=400}
-----------------------------------------
```{r}
renderPlotly({
  data <- gapminder %>%
    filter(
      continent %in% input$continent,
      year >= input$year_range[1] & year <= input$year_range[2],
      country %in% input$country
    )
  
  p <- ggplot(data, aes(x = year, y = .data[[input$metric]], color = country)) +
    geom_line() +
    theme_minimal() +
    labs(title = paste("Trends in", input$metric), y = input$metric, x = "Year")
  
  ggplotly(p)
})
```

Main Tab 3
=======================================================================

Tab 3: Regional Comparison {data-height=400}
--------------------------------------------
```{r}
library(shiny)
library(gapminder)
library(ggplot2)
library(dplyr)


ui <- fluidPage(
  titlePanel("Regional GDP Per Capita Over Time"),
  
  sidebarLayout(
    sidebarPanel(
      selectInput("region", "Select Region", choices = unique(gapminder$continent))
    ),
    
    mainPanel(
      plotOutput("regionalPlot")  
    )
  )
)


server <- function(input, output) {
  

  regionalData <- reactive({
    req(input$region)  
    
    
    filtered_data <- gapminder %>% filter(continent == input$region)
    
    
    validate(
      need(nrow(filtered_data) > 0, "No data available for the selected region.")
    )
    return(filtered_data)
  })
  
  
  output$regionalPlot <- renderPlot({
    data <- regionalData()  
    
    ggplot(data, aes(x = year, y = gdpPercap, color = country)) +
      geom_line() +
      labs(
        title = paste("GDP Per Capita for Region:", input$region),
        x = "Year",
        y = "GDP Per Capita"
      ) +
      theme_minimal()
  })
}


shinyApp(ui = ui, server = server)
```
```{r}
library(shiny)
library(ggplot2)
library(gapminder)

ui <- fluidPage(
  titlePanel("Gapminder Dashboard"),
  
  sidebarLayout(
    sidebarPanel(
      tags$h4("Gapminder Data and Regional Plot"),
      p("Explore the Gapminder data with a regional plot.")
    ),
    
    mainPanel(
      plotOutput("regionalPlot")
    )
  )
)

# Define server logic
server <- function(input, output) {
  
  # Regional Plot - Grouping by Region
  output$regionalPlot <- renderPlot({
    # Check if gapminder data is available
    if (exists("gapminder")) {
      ggplot(gapminder, aes(x = continent, y = lifeExp, color = continent)) +
        geom_boxplot() +
        labs(title = "Life Expectancy by Continent",
             x = "Continent",
             y = "Life Expectancy") +
        theme_minimal()
    } else {
      print("gapminder dataset is not found.")
    }
  })
}

# Run the application
shinyApp(ui = ui, server = server)
```
```{r}
library(shiny)
library(ggplot2)
library(gapminder)

# Define UI
ui <- fluidPage(
  titlePanel("Life Expectancy by Continent"),
  sidebarLayout(
    sidebarPanel(
      h3("Regional Data Plot")
    ),
    mainPanel(
      plotOutput("regionalPlot", height = "500px")
    )
  )
)

# Define server logic
server <- function(input, output) {
  
  # Explicitly load the gapminder dataset
  data(gapminder)  # Ensure the dataset is loaded
  
  output$regionalPlot <- renderPlot({
    
    # Debug messages
    print("Generating regional plot...")
    print(str(gapminder))  # Check the structure of gapminder
    
    # Generate the plot
    ggplot(gapminder, aes(x = continent, y = lifeExp, color = continent)) +
      geom_boxplot() +
      labs(title = "Life Expectancy by Continent",
           x = "Continent",
           y = "Life Expectancy") +
      theme_minimal()
  })
}

# Run the application 
shinyApp(ui = ui, server = server)
```



Social Media and Source Links {data-height=100}
-----------------------------------------------
```{r}
library(shiny)
library(ggplot2)

ui <- fluidPage(
  titlePanel("Social Media Links Dashboard"),
  
  sidebarLayout(
    sidebarPanel(
     
      tags$h4("Follow Gapminder on Social Media:"),
      tags$ul(
        # Facebook link
        tags$li(a(href = "https://www.facebook.com/gapminderfoundation", 
                  target = "_blank", 
                  img(src = "https://upload.wikimedia.org/wikipedia/commons/5/51/Facebook_f_logo_%282019%29.svg", width = "30px", height = "30px"))),
        # Twitter link
        tags$li(a(href = "https://twitter.com/gapminder", 
                  target = "_blank", 
                  img(src = "https://upload.wikimedia.org/wikipedia/commons/6/60/Twitter_Logo_as_of_2021.svg", width = "30px", height = "30px"))),
        # Instagram link
        tags$li(a(href = "https://www.instagram.com/gapminderfoundation", 
                  target = "_blank", 
                  img(src = "https://upload.wikimedia.org/wikipedia/commons/a/a5/Instagram_icon.png", width = "30px", height = "30px"))),
        # YouTube link
        tags$li(a(href = "https://www.youtube.com/user/gapminderfoundation", 
                  target = "_blank", 
                  img(src = "https://upload.wikimedia.org/wikipedia/commons/a/a9/YouTube_icon_%282013-2017%29.svg", width = "30px", height = "30px")))
      )
    ),
    
    mainPanel(
      # Display a paragraph of text or some additional content
      tags$h3("Welcome to Gapminder Social Media Dashboard"),
      tags$p("Click the icons above to follow Gapminder on various social media platforms.")
    )
  )
)


server <- function(input, output) {
  # No plot or data needed, so no server-side processing required.
}


shinyApp(ui = ui, server = server)
```




```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This R Markdown document is made interactive using Shiny. Unlike the more traditional workflow of creating static reports, you can now create documents that allow your readers to change the assumptions underlying your analysis and see the results immediately. 

To learn more, see [Interactive Documents](http://rmarkdown.rstudio.com/authoring_shiny.html).

## Inputs and Outputs

You can embed Shiny inputs and outputs in your document. Outputs are automatically updated whenever inputs change.  This demonstrates how a standard R plot can be made interactive by wrapping it in the Shiny `renderPlot` function. The `selectInput` and `sliderInput` functions create the input widgets used to drive the plot.

```{r eruptions, echo=FALSE}
inputPanel(
  selectInput("n_breaks", label = "Number of bins:",
              choices = c(10, 20, 35, 50), selected = 20),
  
  sliderInput("bw_adjust", label = "Bandwidth adjustment:",
              min = 0.2, max = 2, value = 1, step = 0.2)
)

renderPlot({
  hist(faithful$eruptions, probability = TRUE, breaks = as.numeric(input$n_breaks),
       xlab = "Duration (minutes)", main = "Geyser eruption duration")
  
  dens <- density(faithful$eruptions, adjust = input$bw_adjust)
  lines(dens, col = "blue")
})
```

## Embedded Application

It's also possible to embed an entire Shiny application within an R Markdown document using the `shinyAppDir` function. This example embeds a Shiny application located in another directory:

```{r tabsets, echo=FALSE}
shinyAppDir(
  system.file("examples/06_tabsets", package = "shiny"),
  options = list(
    width = "100%", height = 550
  )
)
```

Note the use of the `height` parameter to determine how much vertical space the embedded application should occupy.

You can also use the `shinyApp` function to define an application inline rather then in an external directory.

In all of R code chunks above the `echo = FALSE` attribute is used. This is to prevent the R code within the chunk from rendering in the document alongside the Shiny components.



